Get Daymet weather data and calculate ET0

### 2023-11-02 Codes/0_1_hpa_counties.Rmd
metadata completed

# Additional counties 
Additional state codes:
these supplement the dryland production data for understanding the impact of drought severity
on dryland crop production

>>>>

# Additional counties 
Additional state codes:
these may supplement the dryland production data for understanding the impact of drought severity
on dryland crop production

### 2023-11-02 Codes/0_2_quickstat.Rmd
metadata completed

```{r process-nass-data}
(
  nass_data_wide <-
    here("Data/data-raw/nass_data_raw.rds") %>%
    readRDS() %>%
    # === drop observations with county_code == 998 ===#
    .[county_code != 998, ] %>%
    # === define state-county code ===#
    .[, sc_code := paste0(state_code, county_code)] %>%
    .[, .(sc_code, year, short_desc, Value)] %>%
    .[, short_desc := case_when(
      str_detect(short_desc, "CORN, GRAIN, IRRIGATED - YIELD") ~ "corn_ir_yield",
      str_detect(short_desc, "CORN, GRAIN, NON-IRRIGATED - YIELD") ~ "corn_nir_yield",
      str_detect(short_desc, "CORN, GRAIN, IRRIGATED - ACRES HARVESTED") ~ "corn_ir_acres",
      str_detect(short_desc, "CORN, GRAIN, NON-IRRIGATED - ACRES HARVESTED") ~ "corn_nir_acres",
      str_detect(short_desc, "SOYBEANS, IRRIGATED - YIELD") ~ "soy_ir_yield",
      str_detect(short_desc, "SOYBEANS, NON-IRRIGATED - YIELD") ~ "soy_nir_yield",
      str_detect(short_desc, "SOYBEANS, IRRIGATED - ACRES HARVESTED") ~ "soy_ir_acres",
      str_detect(short_desc, "SOYBEANS, NON-IRRIGATED - ACRES HARVESTED") ~ "soy_nir_acres"
    )] %>%
    .[, var_type := ifelse(
      str_detect(short_desc, "acres"),
      "acres", "yield"
    )] %>%
    # === irrigated or not? ===#
    .[, ir := ifelse(
      str_detect(short_desc, "nir"),
      "nir", "ir"
    )] %>%
    # === crop type distinction ===#
    .[, crop := ifelse(
      str_detect(short_desc, "corn"),
      "corn", "soy"
    )] %>%
    .[, short_desc := NULL] %>%
    # === long to wide ===#
    dcast(sc_code + year + crop + ir ~ var_type, value.var = "Value")
)
```

>>>>

```{r process-nass-data}
(
  nass_data_wide <-
    here("Data/data-raw/nass_data_raw.rds") %>%
    readRDS() %>%
    # === drop observations with county_code == 998 ===#
    .[county_code != 998, ] %>%
    # === define state-county code ===#
    .[, sc_code := paste0(state_code, county_code)] %>%
    .[, .(sc_code, year, short_desc, Value)] %>%
    .[, short_desc := case_when(
      str_detect(short_desc, "CORN, GRAIN, IRRIGATED - YIELD") ~ "corn_ir_yield",
      str_detect(short_desc, "CORN, GRAIN, NON-IRRIGATED - YIELD") ~ "corn_nir_yield",
      str_detect(short_desc, "CORN, GRAIN, IRRIGATED - ACRES HARVESTED") ~ "corn_ir_acres",
      str_detect(short_desc, "CORN, GRAIN, NON-IRRIGATED - ACRES HARVESTED") ~ "corn_nir_acres",
      str_detect(short_desc, "SOYBEANS, IRRIGATED - YIELD") ~ "soy_ir_yield",
      str_detect(short_desc, "SOYBEANS, NON-IRRIGATED - YIELD") ~ "soy_nir_yield",
      str_detect(short_desc, "SOYBEANS, IRRIGATED - ACRES HARVESTED") ~ "soy_ir_acres",
      str_detect(short_desc, "SOYBEANS, NON-IRRIGATED - ACRES HARVESTED") ~ "soy_nir_acres"
    )] %>%
    .[, var_type := ifelse(
      str_detect(short_desc, "acres"),
      "acres", "yield"
    )] %>%
    # === irrigated or not ===#
    .[, ir := ifelse(
      str_detect(short_desc, "nir"),
      "nir", "ir"
    )] %>%
    # === crop type distinction ===#
    .[, crop := ifelse(
      str_detect(short_desc, "corn"),
      "corn", "soy"
    )] %>%
    .[, short_desc := NULL] %>%
    # === long to wide ===#
    dcast(sc_code + year + crop + ir ~ var_type, value.var = "Value")
)
```

### 2023-11-02 Codes/0_3_get_sat.Rmd
metadata completed

# Preparation
Import saturated thickness data

```{r }
sat_temp <- raster("Data/data-raw/saturated_thickness/SatThick_19851.tif")
```

@@

```{r }
hpa_counties <-
  here("Data/data-processed/base_counties.rds") %>%
  readRDS() %>%
  .[in_hpa == 1, ] %>%
  st_as_sf() %>%
  st_transform(projection(sat_temp))
```

>>>>

# Preparation
Import saturated thickness data to get the CRS.

```{r }
sat_crs <-
  raster("Data/data-raw/saturated_thickness/SatThick_19851.tif") %>%
  raster::projection()
```

@@

```{r }
hpa_counties <-
  here("Data/data-processed/base_counties.rds") %>%
  readRDS() %>%
  .[in_hpa == 1, ] %>%
  st_as_sf() %>%
  st_transform(sat_crs)
```

### 2023-11-02 Codes/0_4_get_weather.Rmd
metadata completed

#--- summarize the data ---#
sum_gm_data <-
  gm_data %>%
  .[month(date) >= 5 & month(date) <= 9, ] %>%
  .[, t_min_max_2 := pmin(pmax((tmin + tmax) / 2 - 10, 0), 20)] %>%
  .[, .(
    # === total balance ===#
    balance = sum(et0 - prcp),
    # === total et0 ===#
    et0 = sum(et0),
    # === total precip ===#
    prcp = sum(prcp),
    # === GDD ===#,
    gdd = sum(t_min_max_2),
    # === EDD ===#
    edd = sum(pmax(0, tmax - 32)),
    # === number of days above x ===#
    days_ab_30 = sum(tmax >= 30),
    # === mean temp ===#
    mean_temp = mean(tmin + tmax) / 2
  ),
  by = .(sc_code, year(date))
  ]

saveRDS(sum_gm_data, here("Data/data-processed", "summrized_gridMET.rds"))
```

>>>>

#--- summarize the data ---#
sum_gm_data <-
  gm_data %>%
  .[month(date) >= 5 & month(date) <= 9, ] %>%
  .[, t_min_max_2 := pmin(pmax((tmin + tmax) / 2 - 10, 0), 20)] %>%
  .[, .(
    # === total balance ===#
    balance = sum(et0 - prcp),
    # === total et0 ===#
    et0 = sum(et0),
    # === total precip ===#
    prcp = sum(prcp)
  ),
  by = .(sc_code, year(date))
  ]

saveRDS(sum_gm_data, here("Data/data-processed", "summarized_gridMET.rds"))
```

### 2023-11-02 Codes/0_6_merge_datasets.rmd
metadata completed

#' Weather
weather_data <- readRDS(here("Data/data-processed/summrized_gridMET.rds"))

@@

```{r}
final_data <-
  merged_data %>%
  #--- yield unit conversion (bu/acre to ton/ha)---#
  .[, yield := yield * 12.553 / 200] %>%
  # === if not in_hpa and irrigated, drop them ===#
  # irrigated, but saturated thickness not observed. These
  # observations cannot be used.
  .[!(in_hpa == 0 & ir == "ir"), ] %>%
  # === if not irrigated, set sat = 0 ===#
  .[ir == "nir", sat := 0] %>%
  # === if sat == 0, but indicated irrigated ===#
  # add 0.1 to distinguish dryland production, where
  # sat is set at 0 above
  .[sat == 0 & ir == "ir", sat := sat + 0.01] %>%
  # === drop if yield or acres are not observed ===#
  .[!is.na(yield) & !is.na(acres), ] %>%
  # === if sat is missing, drop ===#
  #' just 2017 and 2018 data because saturated thickness data
  #' is observed only until 2016
  .[!is.na(sat), ] %>%
  #--- indicator of whether dryland (rainfed) production or not ---#
  .[, dry_or_not := fifelse(sat == 0, 1, 0)] %>%
  #--- sc_code and dry_or_not (used for FE regression) ---#
  .[, sc_dry := paste0(sc_code, "_", dry_or_not)] %>%
  #--- state-year combination ---#
  .[, state_year := paste0(state_name, "_", year)] %>%
  # === number of observations by state-county code ===#
  .[, num_obs := .N, by = sc_dry] %>%
  # .[, num_obs := .N, by = sc_code] %>%
  # === keep only if more than 10 observations are available ===#
  .[num_obs >= 10, ] %>%
  # === order (not necessary) ===#
  .[order(sc_code, year), ]
```

>>>>

#' Weather
weather_data <- readRDS(here("Data/data-processed/summarized_gridMET.rds"))

@@

```{r}
final_data <-
  merged_data %>%
  #--- yield unit conversion (bu/acre to ton/ha)---#
  .[, yield := yield * 12.553 / 200] %>%
  # === if not in_hpa and irrigated, drop them ===#
  # irrigated, but saturated thickness not observed. These
  # observations cannot be used.
  .[!(in_hpa == 0 & ir == "ir"), ] %>%
  #--- drop if irrigated and year > 2016 ---#
  # drops irrigated produciton observations that are missing sat (2017, 2018 fow which sat is not observed). This will keep all the dryland production observations beyond 2016 because sat does not have to be observed for the dryland production.
  .[!(ir == "ir" & year > 2016), ] %>%
  # === drop if both yield and acres are not observed ===#
  .[!is.na(yield) & !is.na(acres), ] %>%
  #--- set sat = NA if not irrigated ---#
  .[ir == "nir", sat := NA] %>%
  #--- sc_code and irrigation status (used for FE regression) ---#
  .[, sc_dry := paste0(sc_code, "_", ir)] %>%
  #--- state-year combination ---#
  .[, state_year := paste0(state_name, "_", year)] %>%
  # === number of observations by state-county code ===#
  .[, num_obs := .N, by = sc_dry] %>%
  # === keep only if more than 10 observations are available ===#
  .[num_obs >= 10, ] %>%
  .[, num_obs := NULL] %>%
  # === order (not necessary) ===#
  .[order(sc_code, year), ] %>%
  dplyr::relocate(sc_code, year)
```
